ManuscriptPURFits

### 2024-02-13 Code/ManuscriptPURFits.Rmd

# Nest all data before photoperiods with no growths were filtered out
NoFilter <- PURPARFits_dummy |>
   filter(PURPhotonDose_day <= 2.0e+06) |>
  mutate(Photoperiod_all = Photoperiod) %>% 
  nest(Data_all = -c("Strain","O2_uM"))

 # Join nested df with all data to df containing filtered data 
  PURPARStrainO2NestModels_filtered <- full_join(PURPARStrainO2NestModels_filtered, NoFilter) %>% 
    rename(Data_filtered = 'data')

@@

# # model with dummy dark, 0 PAR,  growth rates 
# PURPARStrainO2WLNestModels <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )
# 
# 
# # model with dummy dark, 0 PAR,  growth rates
# PURPARStrainO2PeriodWLNest <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "Photoperiod", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )

UnnestTest <- PURPARStrainO2NestModels_filtered %>% 
  # unnest(Data_all, names_sep = "_", keep_empty = TRUE) %>% 
  unnest(GRCplatt_Predict, keep_empty = TRUE)

@@

# PURPARStrainO2NestModels |>
#   filter(!map_lgl(GRC_Model, is.null)) |> #https://stackoverflow.com/questions/57336319/how-to-filter-out-null-elements-of-tibbles-list-column

PURPARStrainO2NestModels_filtered|>
  filter(!map_lgl(GRCplatt_Model, is.null)) |>
  unnest(GRCplatt_Predict, Data_filtered) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = Photoperiod)) + 
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) + 
  labs(title = "PURPARStrainO2NestModels_filtered Platt") +
  theme_bw()

@@

# O2 nested data frame containing the single fit of pooled photoperiods 
 StrainO2filtered_PlattLmModel <- StrainO2_AllModels|>
  # select(-c(data)) |>
  pivot_longer(cols = -c(Strain, O2_uM, Data_filtered, Data_all)) |>
  drop_na(value) |>
  filter(str_detect(name, pattern = "Model") | str_detect(name, pattern = "Glance") | str_detect(name, pattern = "Param") | str_detect(name, pattern = "Predict")) |>
  separate(col = name , into = c("ModelType", "Output"), sep = "_") |>
  pivot_wider(names_from = Output, values_from = value) |>
  filter(!map_lgl(Model, is.null)) |>
  filter(!map_lgl(Glance, is.null)) |>
  mutate(AIC = map_dbl(Glance, possibly(~ as.numeric(pluck(.x, "AIC"))))) |>
  group_by(Strain, O2_uM) |>
  filter(AIC == min(AIC)) |>
  ungroup() |>
  select(-c(Glance))|>
  unnest(Param) |>
  select(-c(statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate, std.error, p.value))

@@

# O2 nested data frame containing the single fit of pooled photoperiods for each strain
StrainO2filtered_PlattLmModel |>
  unnest(Predict, Data_filtered) |>
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL= factor(WL, levels = c("WW", 660, 450)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4"))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL))) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
   geom_text(aes(x = 500000, y = 0.75, label = paste(ModelType), show.legend = FALSE)
  ) +
     scale_colour_manual(values = MCMIXColours) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) +
    labs(title = "Single fit for each strain in each [O2]- all photoperiods included that had growth") +
  theme_bw()

@@

# make wider to join with DF that contains the 1 fit through all photoperiods
Wider_Photoperiods <- StrainO2Photoperiod_PlattLmModel %>% 
  select(Strain, Photoperiod, O2_uM, Model, Predict, data) %>% 
  pivot_wider(names_from = Photoperiod, values_from = c(Model, Predict, data))

@@

# this contains 1 fit through all photoperiods, the photoperiods with no growth were filtered out.
AllPhotoperiods <- StrainO2filtered_PlattLmModel %>% 
  select(Strain, O2_uM, Model, Predict, Data_filtered, Data_all) %>% 
 rename(All = 'Model',
        Predict_all = 'Predict')

@@

remove(Wider_Photoperiods)
remove(AllPhotoperiods)
```

# Unnest predict df for each photoperiod in order to plot fitted 
```{r}
Unnest_16 <- AllFitsCombined %>% 
  select(c(Strain, O2_uM, Predict_16)) %>%  
  unnest(Predict_16, keep_empty = TRUE) %>% 
  mutate(FitCategory = "Photoperiod16")

Unnest_12 <- AllFitsCombined %>% 
  select(c(Strain, O2_uM, Predict_12)) %>%  
  unnest(Predict_12, keep_empty = TRUE) %>% 
  mutate(FitCategory = "Photoperiod12")

Unnest_8 <- AllFitsCombined %>% 
  select(c(Strain, O2_uM, Predict_8)) %>%  
  unnest(Predict_8, keep_empty = TRUE) %>% 
  mutate(FitCategory = "Photoperiod8")

Unnest_4 <- AllFitsCombined %>% 
  select(c(Strain, O2_uM, Predict_4)) %>%  
  unnest(Predict_4, keep_empty = TRUE) %>% 
  mutate(FitCategory = "Photoperiod4")

Unnest_all <- AllFitsCombined %>% 
  select(c(Strain, O2_uM, Predict_all, Data_all)) %>% 
  # unnest(Data_all, keep_empty = TRUE) %>% 
  unnest(Predict_all, keep_empty = TRUE) %>% 
  mutate(FitCategory = "Pooled")

ForFitPlots <- bind_rows(Unnest_all, Unnest_16, Unnest_12, Unnest_8, Unnest_4)

remove(Unnest_16, Unnest_12, Unnest_8, Unnest_4, Unnest_all)

#Mirelle left off here,
# Need to use the DF with all data (all the photoperiods that showed complete no growth )
#next add fits that were sig different, do this dynamically
```{r Plot fits, fig.height= 20, fig.width= 20 }

@@

ForFitPlots |>

@@

    unnest(cols = c(Data_all), names_sep = "_", keep_empty = TRUE) |>
  # unnest(Predict_4, names_sep = "_", keep_empty = TRUE) |>
  # unnest(Predict_8, names_sep = "_", keep_empty = TRUE) |>
  # unnest(Predict_12, names_sep = "_", keep_empty = TRUE) |>
  # unnest(Predict_16, names_sep = "_", keep_empty = TRUE) |>

@@

         Photoperiod = factor(Data_all_Photoperiod, levels = c("16", "12", "8", "4")),

@@

   geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_ReplicateMean_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(Data_all_WL), fill = as.factor(Data_all_WL)), size = 6, alpha = 0.8) +
  geom_point(aes(x = Data_all_PURPhotonDose_day, y = Data_all_deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(Data_all_WL), fill = as.factor(Data_all_WL)), size = 2, alpha = 1.0) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted), data = . %>% filter(FitCategory == "Pooled"), size = 1) + 
  geom_line(aes(x = PURPhotonDose_day, y = .fitted, linetype = "twodash"), data = . %>% filter(FitCategory == "Photoperiod16" & O2_uM == 25), size = 0.5) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted, linetype = "dotdash"), data = . %>% filter(FitCategory == "Photoperiod8"), size = 0.5) +
  geom_line(aes(x = PURPhotonDose_day, y = .fitted,  linetype = "aa"), data = . %>% filter(FitCategory == "Photoperiod4"), size = 0.5) +
  # geom_line(aes(x = PURPhotonDose_day, y = .fitted, shape = as.factor(Photoperiod), linetype = as.factor(Photoperiod)), data = . %>% filter(FitCategory == "Photoperiod4"), size = 0.5) +

@@

   scale_linetype_manual(values = c("solid", "twodash", "dotted","dotdash", "aa"),labels = c("Pooled", "16 hr", "12 hr", "8 hr", "4 hr")) +

@@

 guides(linetype = guide_legend(nrow = 5),

@@

        legend.title =  element_text(size = 25),  #change legend title font size
        legend.text = element_text(size = 25),   #change legend text font size

@@

  unnest(cols = c(Data_all, Predict_all), names_sep = "_", keep_empty = TRUE) |>

@@

  geom_line(aes(x = Predict_all_PURPhotonDose_day, y = Predict_all_.fitted), size = 1) + 

@@

# this contains 1 fit through all photoperiods, the photoperiods with no growth were filtered out.
AllPhotoperiods <- StrainO2filteredPhotoperiod_PlattLmModel %>% 
  select(Strain, O2_uM, Model, Predict) %>% 
 rename(All = 'Model',
        Predict_all = 'Predict')

>>>>

# nest unfiltered data and join with filtered data to have all data to plot later 
AllData <- PURPARFits_dummy |>
   filter(PURPhotonDose_day <= 2.0e+06) |>
    nest(Data_all = -c("Strain","O2_uM"))

PURPARStrainO2NestModels_filtered <- full_join(PURPARStrainO2NestModels_filtered, AllData)

@@

# # model with dummy dark, 0 PAR,  growth rates 
# PURPARStrainO2WLNestModels <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )
# 
# 
# # model with dummy dark, 0 PAR,  growth rates
# PURPARStrainO2PeriodWLNest <- PURPARFits_dummy |>
#   nest(data = -c("Strain","O2_uM", "Photoperiod", "WL")) |>
#   mutate(GRCplatt_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ grcplatt(I = PURPhotonDose_day, a, b, Pmax),
#                               data = .x,
#                               start = list(a = 1e-7, b = 1e-8, Pmax = 1),
#                               lower = c(0, 0, 0),
#                               upper = c(1,1,100),
#                               control = list(maxiter = 500)
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRCplatt_Param = map(GRCplatt_Model, tidy),
#          GRCplatt_Predict = map(GRCplatt_Model, possibly(augment)),
#          GRCplatt_Glance = map(GRCplatt_Model, possibly(glance))
#          )|>
#   mutate(GRClm_Model = map(data, possibly(~nlsLM(deltaOD_Lmu_corr_day ~ LinEqn(x = PURPhotonDose_day,  M, Intercept),
#                               data = .x,
#                               start = LinEqnStart
#                          ), NULL)
#                          )
#          ) |>
#   mutate(GRClm_Param = map(GRClm_Model, tidy),
#          GRClm_Predict = map(GRClm_Model, possibly(augment)),
#          GRClm_Glance = map(GRClm_Model, possibly(glance))
#          )
remove(AllData)

@@

# PURPARStrainO2NestModels |>
#   filter(!map_lgl(GRC_Model, is.null)) |> #https://stackoverflow.com/questions/57336319/how-to-filter-out-null-elements-of-tibbles-list-column 
PURPARStrainO2NestModels_filtered|>
  filter(!map_lgl(GRCplatt_Model, is.null)) |>
  unnest(GRCplatt_Predict, data) |>
   mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c(16, 12, 8, 4))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, colour = Photoperiod)) + 
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) + 
  labs(title = "PURPARStrainO2NestModels_filtered Platt") +
  theme_bw()

@@

# O2 nested data frame containing the single fit of pooled photoperiods 
 StrainO2filtered_PlattLmModel <- StrainO2_AllModels|>
  # select(-c(data)) |>
  pivot_longer(cols = -c(Strain, O2_uM, data, Data_all)) |>
  drop_na(value) |>
  filter(str_detect(name, pattern = "Model") | str_detect(name, pattern = "Glance") | str_detect(name, pattern = "Param") | str_detect(name, pattern = "Predict")) |>
  separate(col = name , into = c("ModelType", "Output"), sep = "_") |>
  pivot_wider(names_from = Output, values_from = value) |>
  filter(!map_lgl(Model, is.null)) |>
  filter(!map_lgl(Glance, is.null)) |>
  mutate(AIC = map_dbl(Glance, possibly(~ as.numeric(pluck(.x, "AIC"))))) |>
  group_by(Strain, O2_uM) |>
  filter(AIC == min(AIC)) |>
  ungroup() |>
  select(-c(Glance))|>
  unnest(Param) |>
  select(-c(statistic)) |>
  pivot_wider(names_from = term, values_from = c(estimate, std.error, p.value))

@@

# O2 nested data frame containing the single fit of pooled photoperiods for each strain
StrainO2filtered_PlattLmModel |>
  unnest(Predict, data) |>
  mutate(O2_uM = factor(O2_uM, levels = c(250, 25, 2.5)),
         WL= factor(WL, levels = c("WW", 660, 450)),
         Strain = factor(Strain, levels = c("MED4", "SS120", "MIT9313")),
         Photoperiod = factor(Photoperiod, levels = c("16", "12", "8", "4"))) |>
  ggplot() +
  geom_point(aes(x = PURPhotonDose_day, y = deltaOD_Lmu_corr_day, shape = as.factor(Photoperiod), color = as.factor(WL), fill = as.factor(WL))) +
  geom_line(aes(x = PURPhotonDose_day, y = `.fitted`)) + 
   geom_text(aes(x = 500000, y = 0.75, label = paste(ModelType), show.legend = FALSE)
  ) +
     scale_colour_manual(values = MCMIXColours) +
   scale_shape_manual(values = PhotoperiodShapes)  +
  facet_grid(rows = vars(O2_uM), cols = vars(Strain)) +
    labs(title = "Single fit for each strain in each [O2]- all photoperiods included that had growth") +
  theme_bw()

@@

@@

# this contains 1 fit through all photoperiods, the photoperiods with no growth were filtered out.
AllPhotoperiods <- StrainO2filtered_PlattLmModel %>% 
  select(Strain, O2_uM, Model, Predict, Data_all, data) %>% 
 rename(All = 'Model',
        Predict_pooled = 'Predict',
        Data_filtered = 'data')

@@

remove(Wider_Photoperiods)
remove(AllPhotoperiods)

AllFitsTest <- AllFitsCombined %>%
  # unnest(Data_all) %>% 
  tidyr::unnest(Predict_4, names_sep = "_", keep_empty = TRUE) %>% 
  tidyr::unnest(Predict_8, names_sep = "_", keep_empty = TRUE) %>%
  tidyr::unnest(Predict_12, names_sep = "_", keep_empty = TRUE) %>%
  tidyr::unnest(Predict_16, names_sep = "_", keep_empty = TRUE) %>%

  rename(Four_PURPhotonDose_day = PURPhotonDose_day,
         Four_deltaOD_Lmu_corr_day = deltaOD_Lmu_corr_day,
         Four_Fitted = .fitted) %>% 
  select(-c(.resid)) %>%
  unnest(Predict_8) %>% 
  rename(Eight_PURPhotonDose_day = PURPhotonDose_day,
         Eight_deltaOD_Lmu_corr_day = deltaOD_Lmu_corr_day,
         Eight_Fitted = .fitted) %>% 
  select(-c(.resid))

```
#Mirelle left off here,
# Need to use the DF with all data (all the photoperiods that showed complete no growth )
#next add fits that were sig different, do this dynamically
```{r Plot fits, fig.height= 20, fig.width= 20 }

@@

@@

  unnest(cols = c(Data_all, Predict_pooled), names_sep = "_", keep_empty = TRUE) |>
  unnest(Predict_4, names_sep = "_", keep_empty = TRUE) |>
  unnest(Predict_8, names_sep = "_", keep_empty = TRUE) |>
  unnest(Predict_12, names_sep = "_", keep_empty = TRUE) |>
  unnest(Predict_16, names_sep = "_", keep_empty = TRUE) |>

@@

@@

  geom_line(aes(x = Predict_pooled_PURPhotonDose_day, y = Predict_pooled_.fitted), size = 1) + 
  geom_line(aes(x = Predict_4_PURPhotonDose_day, y = Predict_4_.fitted, shape = as.factor(Predict_4_Photoperiod), linetype = as.factor(Predict_4_Photoperiod)), size = 0.5) +
  geom_line(aes(x = Predict_8_PURPhotonDose_day, y = Predict_8_.fitted, shape = as.factor(Predict_8_Photoperiod), linetype = as.factor(Predict_8_Photoperiod)), size = 0.5) +
  geom_line(aes(x = Predict_12_PURPhotonDose_day, y = Predict_12_.fitted, shape = as.factor(Predict_12_Photoperiod), linetype = as.factor(Predict_12_Photoperiod)), size = 0.5) +
  geom_line(aes(x = Predict_16_PURPhotonDose_day, y = Predict_16_.fitted, shape = as.factor(Predict_16_Photoperiod), linetype = as.factor(Predict_16_Photoperiod)), size = 0.5) +

@@

@@

@@

@@

  unnest(cols = c(Data_all, Predict_pooled), names_sep = "_", keep_empty = TRUE) |>

@@

  geom_line(aes(x = Predict_pooled_PURPhotonDose_day, y = Predict_pooled_.fitted), size = 1) + 

@@

# this contains 1 fit through all photoperiods, the photoperiods with no growth were filtered out.
AllPhotoperiods <- StrainO2filteredPhotoperiod_PlattLmModel %>% 
  select(Strain, O2_uM, Model, Predict) %>% 
 rename(All = 'Model',
        Predict_pooled = 'Predict')
